---
output:
 
  html_notebook: 
    toc: yes
    toc_depth: 10
    number_sections: yes
  html_document:
    toc: yes
    toc_depth: '10'
    df_print: paged
editor_options:
  chunk_output_type: inline
---

# Loading packages

```{r}
library(dplyr) 
library(tidyr) 
library(ggplot2)
```

# Loading data

```{r}
life_expectancy <- read.csv("UNdata.csv")
```

# Taking a look at the first few rows

```{r}
head(life_expectancy) 
```

# Subsetting and reshaping the life expectancy data

```{r}
subdata <- life_expectancy %>% 
  filter(Year=="2000-2005") %>% 
  select(Country.or.Area, Subgroup, Value) %>% 
  spread(Subgroup, Value)

```

# Taking a look at the first few rows

```{r}
head(subdata)
```

# Plotting male and female life expectancy

```{r}
ggplot(subdata, aes(x=Male, y=Female)) + geom_point()
```

# Adding an abline and changing the scale of axes of the previous plots

```{r}
ggplot(subdata, aes(x=Male, y=Female)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, linetype=2) +
  scale_x_continuous(limits=c(35,85)) +
  scale_y_continuous(limits=c(35,85))

ggplot(subdata, aes(x=Male, y=Female))+
  geom_point(colour="white", fill="chartreuse3", shape=21, alpha=.55, size=5)+
  geom_abline(intercept = 0, slope = 1, linetype=2)+
  scale_x_continuous(limits=c(35,85))+
  scale_y_continuous(limits=c(35,85))+
  labs(title="Life Expectancy at Birth by   Country",
       subtitle="Years. Period: 2000-2005. Average.",
       caption="Source: United Nations Statistics Division",
       x="Males",
       y="Females")
```

# Subseting data to obtain countries of interest

```{r}
top_male <- subdata %>% arrange(Male-Female) %>% head(3)
top_female <- subdata %>% arrange(Female-Male) %>% head(3)
```

# Adding text to the previous plot to label countries of interest

```{r}
ggplot(subdata, aes(x=Male, y=Female, label=Country.or.Area))+
  geom_point(colour="white", fill="chartreuse3", shape=21, alpha=.55, size=5)+
  geom_abline(intercept = 0, slope = 1, linetype=2)+
  scale_x_continuous(limits=c(35,85))+
  scale_y_continuous(limits=c(35,85))+
  labs(title="Life Expectancy at Birth by Country",
       subtitle="Years. Period: 2000-2005. Average.",
       caption="Source: United Nations Statistics Division",
       x="Males",
       y="Females")+
  geom_text(data=top_male, size=3)+
  geom_text(data=top_female, size=3)+
  theme_bw()
```

# Subsetting, mutating and reshaping the life expectancy data

```{r}
subdata2 <- life_expectancy %>% 
  filter(Year %in% c("1985-1990", "2000-2005")) %>% 
  mutate(Sub_Year=paste(Subgroup, Year, sep="_")) %>% 
  mutate(Sub_Year=gsub("-", "_", Sub_Year)) %>% 
  select(-Subgroup, -Year) %>% 
  spread(Sub_Year, Value) %>% 
  mutate(diff_Female = Female_2000_2005 - Female_1985_1990, 
         diff_Male   = Male_2000_2005 - Male_1985_1990)
```

# Taking a look at the first few rows

```{r}
head(subdata2)
```

# Doing a nice first version of the plot with abline, scaling axis and adding labels

```{r}
ggplot(subdata2, aes(x=diff_Male, y=diff_Female, label=Country.or.Area))+
  geom_point(colour="white", fill="chartreuse3", shape=21, alpha=.55, size=5)+
  geom_abline(intercept = 0, slope = 1, linetype=2)+
  scale_x_continuous(limits=c(-25,25))+
  scale_y_continuous(limits=c(-25,25))+
  labs(title="Life Expectancy at Birth by Country in Years",
       subtitle="Difference between 1985-1990 and 2000-2005. Average.",
       caption="Source: United Nations Statistics Division",
       x="Males",
       y="Females")+
  theme_bw()
```

# Adding an hline and vline to previous plots

```{r}
ggplot(subdata2, aes(x=diff_Male, y=diff_Female, label=Country.or.Area))+
  geom_point(colour="white", fill="chartreuse3", shape=21, alpha=.55, size=5)+
  geom_abline(intercept = 0, slope = 1, linetype=2)+
  scale_x_continuous(limits=c(-25,25))+
  scale_y_continuous(limits=c(-25,25))+
  geom_hline(yintercept=0, linetype=2)+
  geom_vline(xintercept=0, linetype=2)+
  labs(title="Life Expectancy at Birth by Country",
       subtitle="Years. Difference between 1985-1990 and 2000-2005. Average.",
       caption="Source: United Nations Statistics Division",
       x="Males",
       y="Females")+
  theme_bw()
```

# Subseting data to obtain countries of interest

```{r}
top <- subdata2 %>% arrange(diff_Male+diff_Female) %>% head(3)
bottom <- subdata2 %>% arrange(-(diff_Male+diff_Female)) %>% head(3)
```

# Adding text to the previous plot to label countries of interest

```{r}
ggplot(subdata2, aes(x=diff_Male, y=diff_Female, label=Country.or.Area), guide=FALSE)+
  geom_point(colour="white", fill="chartreuse3", shape=21, alpha=.55, size=5)+
  geom_abline(intercept = 0, slope = 1, linetype=2)+
  scale_x_continuous(limits=c(-25,25))+
  scale_y_continuous(limits=c(-25,25))+
  geom_hline(yintercept=0, linetype=2)+
  geom_vline(xintercept=0, linetype=2)+
  labs(title="Life Expectancy at Birth by Country",
       subtitle="Years. Difference between 1985-1990 and 2000-2005. Average.",
       caption="Source: United Nations Statistics Division",
       x="Males",
       y="Females")+
  geom_text(data=top, size=3)+
  geom_text(data=bottom, size=3)+
  theme_bw()
```
